---
title: "Data Transformation Exercise One - Formative Submission"
author: "LSE Candidate Number: 51050"
date: "Autumn term 2024"
output: html_document
---

## Initial Dataset

The figure below loads the first few values from pre-existing dataset called 'cars' in R. 

``` {r, echo=FALSE}
library(knitr)

#kable() outputs data in tabular format when knitted in HTML
kable(head(cars), caption = "Figure 1. Summary Dataset showing the speed and distance of cars")
```

### Applying data transformation

To transform the dataset, the code below uses 'data' as an argument of a function that assigns each car with a different id. This helps distinguish the observations between the different cars.

#### Code chunk with function applying transformation to dataset

```{r, echo=TRUE, message= FALSE, warning= FALSE}
library(tidyverse)

#labels_rowid function adds a 'row id' column to 
#state the different observations
label_rowid <- function(data){
  transformed_table <- rowid_to_column(data, var = "unique id")
  return(transformed_table)
}
new_cars_data <- kable(head(label_rowid(cars)), caption = "Figure 2. Summary of transformed table of Figure 1 whereby each car is assigned a unique i.d.")
new_cars_data
```

### Applying a second data transformation

For the second function, a transformation to the dataset is applied whereby time is calculated by using the formula $distance \div speed$, and this is added as a new column.

#### Code chunk with function applying second transformation to dataset
<font size="2"> For the function below, I used some assistance from ChatGPT with the prompt "how can I create a function in r that creates a new column in a dataframe which contains values that are a product of the values from the previous two columns, don't give me the direct answer explain the steps to me instead." </font>
```{r, echo=TRUE, message= FALSE, warning= FALSE}

#This function calculates 'time' by dividing last row (distance)
#with the second row (speed) and adds it to a new column
calc_time <- function(data_new){
  distance_col <- data_new[,ncol(data_new)]
  speed_col <- data_new[,ncol(data_new) - 1]
  time <- distance_col / speed_col
  data_new$time <- time
  return(data_new)
}
time_cars_data <- kable(head(calc_time(cars)), caption = "Figure 3. Applying second transformation to the dataset whereby time is calculated and added to a new column")
time_cars_data
```
