---
title: "Data Transformation Exercise One - Formative Submission"
author: "LSE Candidate Number: 51050"
date: "Autumn term 2024"
output: html_document
---

## Initial Dataset

The figure below loads the first few values from pre-existing dataset called 'cars' in R. 

``` {r, echo=FALSE}
library(knitr)

#kable() outputs data in tabular format when knitted in HTML
kable(head(cars), caption = "Figure 1. Summary Dataset showing the speed and distance of cars")
```

### Applying data transformation

To transform the dataset, the code below uses 'data' as an argument of a function that assigns each car with a different id. This helps distinguish the observations between the different cars.

#### Code chunk with function applying transformation to dataset

```{r, echo=TRUE, message= FALSE, warning= FALSE}
library(tidyverse)

#labels_rowid function adds a 'row id' column to 
#state the different observations
label_rowid <- function(data){
  transformed_table <- rowid_to_column(data, var = "unique id")
  return(transformed_table)
}
new_cars_data <- kable(head(label_rowid(cars)), caption = "Figure 2. Summary of transformed table of Figure 1 whereby each car is assigned a unique i.d.")
new_cars_data
```
